---
title: "HW 05 Soultions"
author: "Elijah Willie"
date: "November 08, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this assignment, I will attempt to do the following tasks:

* Writing functions that are useful to analyze the gapminder dataset
* Work with a nested data frame

## Load in required libraries

First we will load in all the libraries we will be using for this assignment.
```{r}
library(gapminder)
suppressMessages(library("tidyverse"))
library(knitr)
suppressMessages(library(reshape2))
suppressMessages(library(Hmisc))
suppressMessages(library(gridExtra))
library(knitr)
suppressMessages(library(MASS))
suppressMessages(library(broom))
```


## Part 1: Writing Functions


### For this section, I will do the following:

* Fit a model model linear model that predicts `life expectancy` for `year`, `pop` or `gdpPercap` for a given country.
* Create a function to plot a linear fit

#### Create a function that fits a linear model

```{r Fit a linear model}
fit_model <- function(Country, variable, data){
#extract the data for the specific country
data_country <-  as.data.frame(data) %>% 
  filter(country == Country)

#fit the model
if(variable == "pop"){
  model_fit <- lm(lifeExp ~ pop, data_country)
}
if(variable == "gdpPercap"){
  model_fit  <- lm(lifeExp ~ gdpPercap, data_country)
}
if(variable == "year"){
  model_fit <- lm(lifeExp ~ year, data_country)
}

return(model_fit)
}

```


#### Create a function that plots the a given regression fit

```{r Function to plot linear fit}
plotReg <- function (fit) {

require(ggplot2)

p <- ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
return(p)
}

```


#### Test out the functions
Predict life expectancy for `Liberia` and `Ghana` using population.

```{r Predict with population}
#create a set of countries
countries <- c("Liberia", "Ghana")

#fit a linear model using population
res <- lapply(countries, fit_model, variable = "pop", data = gapminder)

#generate the plots
plots <- lapply(res, plotReg)

#put them on a grid
grid.arrange(plots[[1]], plots[[2]])
```

Looking at the plots above, we see that population does a much better job at predicting life expectancy when for `Ghana` When compared with `Liberia`.

Now predict again, but this time use `gdpPercap`

```{r Predict with gdpPercap} 
#fit a linear model using gdpPercap
res <- lapply(countries, fit_model, variable = "gdpPercap", data = gapminder)

#generate the plots
plots <- lapply(res, plotReg)

#put them on a grid
grid.arrange(plots[[1]], plots[[2]])

```

Here we see that `gdpPercap` does a pretty terrible job at predicting life xpectancy. Maybe using year would help.

```{r Predict with Year}
#fit a linear model using population
res <- lapply(countries, fit_model, variable = "year", data = gapminder)

#generate the plots
plots <- lapply(res, plotReg)

#put them on a grid
grid.arrange(plots[[1]], plots[[2]])

```

That is much better indeed. We still see that it performs much better for `Ghana` than `Liberia`.


## Part 2 Work with a nested data frame

In the part, I will attempt to do the following

* Nest the data by country (and continent).
* Fit a model of life expectancy against year. Possibly quadratic, possibly robust.
* Use functions for working with fitted models or the `broom` package to get information out of your linear models.
* Use the usual dplyr, tidyr, and ggplot2 workflows to explore, e.g., the estimated cofficients.
* I will be following the tutorial [here](http://stat545.com/block024_group-nest-split-map.html) closely, but with my own twists and turns.

```{r nesting}
#nest the gapminder dataset by country and continent
gap_nested <- gapminder %>% 
   group_by(continent, country) %>% 
   nest()

#lets get some information on the new dataset
gap_nested
nrow(gap_nested)
ncol(gap_nested)
```

We see that the new dataframe has $142$ rows and $3$ columns

Now lets take a look at some of the elements of the new nested dataframe

```{r Inspect Dataframe}
#we can use indices to get information about the data
kable(gap_nested[[1, "data"]])

#let see if we can get the country
gap_nested$country[[1]]

#we can even get data for a specific country
gap_nested[gap_nested$country == "Liberia",][["data"]]

#lets try it for another country
gap_nested[gap_nested$country == "Ghana",][["data"]]
```



Now lets fit some  models to the data and see how they behave
```{r}
#fit a linear model
lin_fit <- lm(lifeExp ~ year, data = gap_nested[[1, "data"]])
paste("linear fit summary: ")
summary(lin_fit)
#compute the AIC and BIC for this model
paste("linear fit AIC and BIC are: ", AIC(lin_fit), BIC(lin_fit))

#fit a quadratic
quad_fit <- lm(lifeExp ~ poly(year, 2), data = gap_nested[[1, "data"]])
paste("Quad fit summary: ")
summary(quad_fit)
#compute the AIC and BIC for this model
paste("Quad fit AIC and BIC are: ", AIC(lin_fit), BIC(lin_fit))

#fit a robust regression
#note that we will need to use the MASS library to do this
robust_fit <- rlm(lifeExp ~ year, data = gap_nested[[1, "data"]])
paste("Robust fit summary: ")
summary(robust_fit)
#compute the AIC and BIC for this model
paste("Robust fit AIC and BIC are: ",AIC(robust_fit), BIC(robust_fit))
```

We see that for `lifeEXp`, there is no difference in using a `linear`, `quadratic`, or `robust` model. They all return the same `AIC` and `BIC` values.


Now all this seems cumbesome, so lets create another function to fit the linear and the robust for us. :relieved:

```{r nested fit function}
nested_fit <- function(Country, type = "linear", variable = "pop"){
  country_dat <- as.data.frame(gap_nested[gap_nested$country == Country,][["data"]])
  if(type == "robust"){
    if(variable == "pop") fit <- rlm(lifeExp ~ pop, data = country_dat)
    if(variable == "year") fit <- rlm(lifeExp ~ year, data = country_dat)
    if(variable == "gdpPercap") fit <- rlm(lifeExp ~ gdpPercap, data = country_dat)
  }
  if(type == "linear"){
    if(variable == "pop") fit <- lm(lifeExp ~ pop, data = country_dat)
    if(variable == "year") fit <- lm(lifeExp ~ year, data = country_dat)
    if(variable == "gdpPercap") fit <- lm(lifeExp ~ gdpPercap, data = country_dat)
  }
  return(fit)
}
```

Lets test this function out. :smirk:
```{r}
#compute a fit for liberia using a regular linear model
res <- nested_fit("Liberia", variable = "pop")
#summarize the model
summary(res)
#compute AIC and BIC
#using the function defined earlier, plot this fit
plotReg(res)

#now fit a robust model
res_rob <- nested_fit("Liberia", type = "robust", variable = "pop")

#show some summary
summary(res_rob)
```

That was fun, now lets use the `map` function to do this for a few countries from part $1$. :wink:

```{r Map to few countries}
#use the map function to fit to two countries using year as predictor
fits <- map(countries,nested_fit)
fits

```

We can finally do this for all countries in our dataset. :muscle:

```{r Map to all countries}
#apply this function to all countries in the dataset
gap_nested <- gap_nested %>% 
   mutate(fit = map(country, nested_fit))

#show the result
gap_nested
```

Finally, we will use the `broom` library to tidy up the results

```{r tidy the data}
#apply broom to each country
gap_nested <- gap_nested %>% 
  mutate(tidy = map(fit, tidy))

#have a look at the result
gap_nested

#lets have a look at the tidy table for one of the countries
gap_nested[gap_nested$country == "Liberia",][["tidy"]]

#we can finally return to the original data type for which we started using all the information we have obtained
gap_coefs <- gap_nested %>% 
   dplyr::select(continent, country, tidy) %>% 
   unnest(tidy)
#show the final table
gap_coefs
```

Now isn't that beautiful!! :smiley:
